---
title: "Doubling Rate: Deaths and Confirmed Cases"
output: html_notebook
---

# Theory
Calculating the doubling rate based on the prior 7 days:

$t_2 - t_1 = 7$

$T_d = (t_2 - t_2)\frac{ln(2)}{ln(\frac{q2}{q1})}$

Example of Surat city Table 1 shows data wise cumulative COVID19 cas-es  reported  in  Surat  city.3  According  to  Formula2using  the  data  of  COVID19  Doubling  rate/time  calculated  for  24th   March   2020   considering   17thMarch 2020 is day 1 ( t1 = 1), 24th March 2020 is Day 8  (t2 =  8)  ,  q1 is  number  of  cases  on  day  1  (17thMarch 2020 ) = 1 ,q2 number of cases on day 8 (24thMarch  2020)  =5  ,  ln(2)  is  natural  logarithmic  value  of 2 . 

source: [Doubling Time and its Interpretation for COVID 19 Cases](http://njcmindia.org/uploads/11-3_141-143.pdf)


## Method
1. Get confirmed/death data
2. Calculate doubling rate for each entry on a 7-day basis (creating a column with doubling rates)
3. plot

# Importing
```{r}
library(tidyverse)
library(lubridate) # dates
library(ggplot2)
```


# Reading data
```{r}
# provincial_total_confirmed <- read.csv("data/covid19za_provincial_cumulative_timeline_confirmed.csv")
owid_sa_data <- read.csv("data/owid-covid-data_SouthAfricaExtracted_big.csv")
# death_data <- read.csv("data/covid19za_provincial_cumulative_timeline_deaths.csv")
head(provincial_total_confirmed)
```


## Calculate doubling rate
```{r doubling rate formula}
# calculate doubling rate formula
calculate_doubling_rate <- function(num_days, q1, q2) {
  doubling_rate <- num_days*(log(2)/(log(q2/q1)))
  return(doubling_rate)
}
```

```{r confirmed and death doubling rate}
  # Get total confirmed/death cases per date
  confirmed_deaths_doubling <- owid_sa_data %>% 
  filter(!is.na(total_cases)) %>% 
  select(date, total_cases, total_deaths) %>% 
  mutate(date = as.Date(date))
  
  # create subset of confirmed/death to apply doubling rate function to
  num_days <- 7
  after_num_days <- confirmed_deaths_doubling %>% tail(-num_days)
  
  # add num_days temp rows at end
  date <- rep(c(Sys.Date()), num_days)
  total_cases <- rep(c(0), num_days)
  total_deaths <- rep(c(0), num_days)
  data_temp <- data.frame(date, total_cases, total_deaths)
  data_temp <- as.data.frame(data_temp)
  after_num_days <- rbind(after_num_days, data_temp, make.row.names=FALSE)
  
  # calculate confirmed doubling rate
  confirmed_deaths_doubling$confirmed_doubling_rate <- with(confirmed_deaths_doubling, calculate_doubling_rate(num_days, confirmed_deaths_doubling$total_cases, after_num_days$total_cases))
  # calculate deaths doubling rate
  confirmed_deaths_doubling$deaths_doubling_rate <- with(confirmed_deaths_doubling, calculate_doubling_rate(num_days, confirmed_deaths_doubling$total_deaths, after_num_days$total_deaths))
  # remove num_days 0s at end
  confirmed_deaths_doubling <- confirmed_deaths_doubling %>% head(-num_days)
  head(confirmed_deaths_doubling)
```


```{r}
ggplot(confirmed_deaths_doubling, aes(date)) + 
  geom_line(aes(y = confirmed_doubling_rate), color = theology_purple, size=1) +
  geom_line(aes(y = deaths_doubling_rate), color = law_winered, size=1) +
  labs(title= "Confirmed doubling rate",
       caption = "Updated: June 2021",
       x = NULL,
       y = NULL) +
  theme_minimal()
```


<!-- Draft -->
<!-- ```{r confirmed cases doubling rate} -->
<!--   # Get total confirmed cases per date -->
<!--   confirmed_doubling <- provincial_total_confirmed %>%  -->
<!--   filter(!is.na(total_cases)) %>%  -->
<!--   select(date, total_cases) %>%  -->
<!--   mutate(date = as.Date(date)) -->

<!--   # create subset of confirmed to apply doubling rate function to -->
<!--   num_days <- 7 -->
<!--   confirmed_after_num_days <- confirmed_doubling %>% tail(-num_days) -->

<!--   # add num_days 0s at end -->
<!--   date <- rep(c(Sys.Date()), num_days) -->
<!--   total_cases <- rep(c(0), num_days) -->
<!--   data_temp <- data.frame(date, total_cases) -->
<!--   data_temp <- as.data.frame(data_temp) -->
<!--   confirmed_after_num_days <- rbind(confirmed_after_num_days, data_temp, make.row.names=FALSE) -->

<!--   # remove num_days 0s at end -->
<!--   confirmed_doubling$confirmed_doubling_rate <- with(confirmed_doubling, calculate_doubling_rate(num_days, confirmed_doubling$total_cases, confirmed_after_num_days$total_cases)) -->
<!--   confirmed_doubling <- confirmed_doubling %>% head(-num_days) -->
<!--   head(confirmed_doubling) -->
<!-- ``` -->

<!-- ```{r deaths doubling rate} -->
<!--   # Get total deaths per date -->
<!--   deaths_doubling <- death_data %>% -->
<!--   mutate(date = dmy(date)) %>% -->
<!--   mutate(total_deaths = total) %>% -->
<!--   select(date, total_deaths) -->

<!--   # create subset of deaths to apply doubling rate function to -->
<!--   num_days <- 7 -->
<!--   deaths_after_num_days <- deaths_doubling %>% tail(-num_days) -->

<!--   # add num_days 0s at end -->
<!--   date <- rep(c(Sys.Date()), num_days) -->
<!--   total_deaths <- rep(c(0), num_days) -->
<!--   data_temp <- data.frame(date, total_deaths) -->
<!--   data_temp <- as.data.frame(data_temp) -->
<!--   deaths_after_num_days <- rbind(deaths_after_num_days, data_temp, make.row.names=FALSE) -->

<!--   # remove num_days 0s at end -->
<!--   deaths_doubling$deaths_doubling_rate <- with(deaths_doubling, calculate_doubling_rate(num_days, deaths_doubling$total_deaths, deaths_after_num_days$total_deaths)) -->
<!--   deaths_doubling <- deaths_doubling %>% head(-num_days) -->
<!--   head(deaths_doubling) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- confirmed_deaths_doubling <- confirmed_doubling %>%  -->
<!--   inner_join(deaths_doubling, by = "date") -->
<!-- head(confirmed_deaths_doubling) -->
<!-- ``` -->



