---
title: "Topic Modelling"
output: html_notebook
---

<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Setup -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->
# Setup
```{r set colours}
# Microsoft Sharepoint main colours
# https://docs.microsoft.com/en-us/sharepoint/dev/design/themes-colors
ms_red <- "#A4262C"
ms_orange <- "#CA5010"
ms_gold <- "#8F7034"
ms_green <- "#407855"
ms_teal <- "#038387"
ms_blue <- "#0078d4"
ms_darkblue <- "#40587C"
ms_indigo <- "#4052AB"
ms_plum <- "#854085"
ms_purple <- "#8764B8"
ms_coolgrey <- "#737373"
ms_warmgrey <- "#867365"

ms_colors <- c(ms_darkblue, ms_red, ms_orange, ms_teal, ms_gold, ms_coolgrey, ms_purple, ms_warmgrey, ms_green, ms_blue, ms_indigo, ms_plum)
```

```{r custom themes}
custom_column_basic_theme <- function () { 
    theme_minimal(base_size=12, base_family="Avenir") +
        theme(
            panel.background  = element_blank(),
            panel.grid.minor = element_blank(),
            legend.position="none", # no legend
            plot.title = element_text(size = 12, hjust = 0),
            plot.subtitle = element_text(size = 10, hjust = 0),
            axis.title.y = element_text(size = 10),
            axis.title.x = element_text(size = 10),
            axis.text.x = element_text(size = 6)
        )
}
```

```{r import libraries}
library(tidytext)
library(lubridate)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(topicmodels) # topic modelling
library(ldatuning) # lda tuning for topic modelling
library(reshape2)
library(vader) # sentiment analysis
```
```{r global (static) variables}
TOP_WORDS <- 10 # number of top words to show when plotting topics
TUNING <- FALSE # NOTE: set to TRUE to run ldatuning

# LDA tuning number of topic variables
NUM_TOPICS_3 <- 3
NUM_TOPICS_4 <- 4 # all outlet topics per month have this number of topics
NUM_TOPICS_5 <- 5
NUM_TOPICS_6 <- 6

# sentiment analysis variables
SENTIMENT_WORDS_PER_TOPIC <- 200

# media outlet string names
DM_NAME <- "Daily Maverick"
SABC_NAME <- "SABC News"
EWN_NAME <- "Eyewitness News"
N24_NAME <- "News24"
OUTLETS_TO_PLOT <- c(DM_NAME, SABC_NAME, EWN_NAME, N24_NAME)

NYT_NAME <- "The New York Times"
ECON_NAME <- "The Economist"
BBC_NAME <- "BBC News (World)"
AP_NAME <- "The Associated Press"
GLOBAL_OUTLETS_TO_PLOT <- c(NYT_NAME, ECON_NAME, BBC_NAME, AP_NAME)

# plotting variables
SA_SUFFIX <- "South Africa"
GLOBAL_SUFFIX <- "Global"
MONTHS_TO_PLOT <- seq(as.Date("2021/1/1"), by = "month", length.out = 6)
MONTH_LIST <- c("2021-01-01", "2021-02-01", "2021-03-01", "2021-04-01", "2021-05-01", "2021-06-01")
```



```{r extract relevant data from read in csv}
extract_relevant_data <- function(df) {
  return(df %>% select(status_id, name, created_at, text))
}
```

```{r read data}
# SA media organisations
SABCNews_tweets <- extract_relevant_data(read.csv("data/sabcnews.csv"))
DailyMav_tweets <- extract_relevant_data(read.csv("data/dailymaverick.csv"))
EWN_tweets <- extract_relevant_data(read.csv("data/ewnupdates.csv"))
News24_tweets <- extract_relevant_data(read.csv("data/news24.csv"))
SA_media_outlet_list <- list(DailyMav_tweets, SABCNews_tweets, EWN_tweets, News24_tweets)

head(SA_media_outlet_list)


# Global
nyTimes_tweets <- extract_relevant_data(read.csv("data/nytimes.csv"))
economist_tweets <- extract_relevant_data(read.csv("data/theeconomist.csv"))
bbcworld_tweets <- extract_relevant_data(read.csv("data/bbcworld.csv"))
AP_tweets <- extract_relevant_data(read.csv("data/ap.csv"))
global_media_outlet_list <- list(nyTimes_tweets, economist_tweets, bbcworld_tweets, AP_tweets)

head(global_media_outlet_list[1])
```
<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Wrangling: Topic modelling -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->
# Data Wrangling: Topic Modelling
## Data Cleaning
```{r extract tweets related to covid and combine covid tweets into a list}
# extract covid tweets function
get_covid_tweets <- function(df) {
relevant_df <-  df[grepl("covid", df[["text"]]) | 
                  grepl("Covid", df[["text"]]) |
                  grepl("corona", df[["text"]]) |
                  grepl("Corona", df[["text"]]) |
                  grepl("Pandemic", df[["text"]]) |
                  grepl("pandemic", df[["text"]]), ]
relevant_df <- relevant_df %>%
  mutate(created_at = as.Date(created_at))
return(relevant_df)
}

# South African: combine relevant (Covid-related) tweets for all media outlets into list
covid_tweets_SA <- list(get_covid_tweets(DailyMav_tweets), 
                        get_covid_tweets(SABCNews_tweets),
                        get_covid_tweets(EWN_tweets),
                        get_covid_tweets(News24_tweets))

# Global: combine relevant (Covid-related) tweets for all media outlets into list
covid_tweets_global <- list(get_covid_tweets(nyTimes_tweets), 
                            get_covid_tweets(AP_tweets),
                            get_covid_tweets(bbcworld_tweets), 
                            get_covid_tweets(economist_tweets))
```

```{r combine covid-related tweets}
# combine all Covid-related  tweets into one dataframe
tweets_SA <- bind_rows(covid_tweets_SA)
tweets_global <- bind_rows(covid_tweets_global)
```

```{r media outlets}
SA_media_outlets <- unique(tweets_SA[c("name")])
global_media_outlets <- unique(tweets_global[c("name")])
```

```{r tidy tweets - remove stopwords and clean data}
clean_tweets <- function(tweets_in) {
  # define custom stopwords
  my_stopwords <- c("covid", "covid19", "corona", "coronavirus", "pandemic", "#sabcnews",
                    "maverick", "amp", "#covid", "#covid19", "#coronavirus",
                    "south", "africa")
  # remove stopwords, links, mentions, special characters, and numbers
  tidy_tweets <- tweets_in %>% 
    unnest_tokens(word, text, token = "tweets") %>%
    filter(!word %in% stop_words$word,
           !word %in% str_remove_all(stop_words$word, "'"), # remove normal english stopwords that would contain ' that was lost by unnesting
           !word %in% my_stopwords,
           !str_starts(word, "http"),
           !str_starts(word, "@"),
           str_detect(word, "^[a-z#]*$")) %>%
    mutate(word = gsub("\\d+", "", word))

  # change "vaccines" to "vaccine" so it is treated as the same word
  tidy_tweets["word"][tidy_tweets["word"] == "vaccines"] <- "vaccine"
  return(tidy_tweets)
}

tidy_tweets_SA <- clean_tweets(tweets_SA)
tidy_tweets_global <- clean_tweets(tweets_global)
```

## Topic Modelling Setup

```{r extract top words per topic}
# topics arg: df containing word, topic and beta (probability that the word occurs in that topic) values
# words_per_topic: number of top words to get
top_words_per_topic <- function(topics, words_per_topic) {
  return(topics %>%
  group_by(topic) %>%
  slice_max(beta, n = words_per_topic) %>% # get most relevant words for that topic
  ungroup() %>%
  arrange(topic, -beta)) # arrange words with decreasing relevance to the topic
}
```


```{r LDA tuning}
# to find optimal number of topics to model for given document term matrix
# source: https://cran.r-project.org/web/packages/ldatuning/vignettes/topics.html
lda_tuning <- function(orgs_dtm) {
  result <- FindTopicsNumber(
      orgs_dtm,
      topics = seq(from = 2, to = 8, by = 1),
      metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
      method = "Gibbs",
      control = list(seed = 77),
      mc.cores = 2L,
      verbose = TRUE
    )
  # choose extreme in following plots as optimal number of topics
  FindTopicsNumber_plot(result)
}
```

```{r extract topics using LDA output}
get_topics <- function(word_counts, nr_topics) {
  # create document term matrix from word count
  orgs_dtm <- word_counts %>%
    cast_dtm(name, word, n)
  
  # checking for ideal number of topics (NOTE: set to TRUE to run ldatuning)
  if (TUNING) {
    lda_tuning(orgs_dtm)
  }
  
  # apply topic modelling to dtm and extract relevant data from LDA output
  orgs_lda <- LDA(orgs_dtm, k = nr_topics, control = list(seed = 1234))
  return(tidy(orgs_lda, matrix = "beta"))
}
```


## Overall topics

```{r apply topic modelling to get overall topics}
overall_topics <- function(tidy_tweets_in, nr_topics) {
  # word count by media organisation
  word_counts <- tidy_tweets_in %>%
    select(name, word, created_at) %>%
    count(name, word, sort = TRUE) %>% # count how often each org uses each word
    ungroup()
  
  orgs_topics_overall <- get_topics(word_counts, nr_topics)
  
  return(orgs_topics_overall)
}
```
 
## Overall Topics for Specific Month
 
```{r apply topic modelling to get topics per month}
monthly_topics <- function(tidy_tweets_in, past_month, nr_topics) {
  # word count by media organisation for specified month
  word_counts <- tidy_tweets_in %>%
    mutate(time_floor = floor_date(created_at, unit = "month")) %>% # create time bins (monthly)
    select(name, word, time_floor) %>%
    filter(time_floor == as.Date(past_month)) %>%
    count(name, word, sort = TRUE) %>% # count how often each org uses each word
    ungroup()
    
  orgs_topics_month <- get_topics(word_counts, nr_topics)
  
  return(orgs_topics_month)
}
```

 
## Topics for Specific Media Outlet

```{r apply topic modelling to get topics per media outlet}
org_topics <- function(tidy_tweets_in, org, nr_topics) {
  # word count for specified media organisation
  word_counts <- tidy_tweets_in %>%
    select(name, word) %>%
    filter(name == org) %>%
    count(name, word, sort = TRUE) %>% # count how often the org uses each word
    ungroup()
    
  org_topics <- get_topics(word_counts, nr_topics)
  
  return(org_topics)
}
```


## Topics for specific Media Outlet for specific Month

```{r apply topic modelling to get topics per organisation per month}
monthly_org_topics <- function(tidy_tweets_in, org, past_month, nr_topics) {
  # word count for specified media organisation for a specific month
  word_counts <- tidy_tweets_in %>%
    mutate(time_floor = floor_date(created_at, unit = "month")) %>%
    select(name, word, time_floor) %>%
    filter(name == org) %>%
    filter(time_floor == as.Date(past_month)) %>%
    count(name, word, sort = TRUE) %>% # count how often the org uses each word
    ungroup()
  
  if (length(word_counts$name) == 0) {
    print(paste(org, "did not tweet the month starting on", past_month))
    return()
  }
    
  org_month_topics <- get_topics(word_counts, nr_topics)
  
  return(org_month_topics)
}
```


<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Visualisation: Topic modelling -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

# Data Visualisation: Topic Modelling

## Overall topics

```{r plot topics}
plot_overall_topics <- function(tidy_tweets_in, nr_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # get the topics to plot
  orgs_topics <- overall_topics(tidy_tweets_in, nr_topics)
  
  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # order terms to decreasing relevance
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS) # force number of terms to be TOP_WORDS per topic
  
  # plot top terms per topic (sorted by beta relevance)
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    ggplot(aes(beta, term, fill = factor(topic))) +
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics between Jan-June 2021"),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # combine plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```

```{r plot overall topics using selected number of topics (lda tuning)}
# South Africa
plot_overall_topics(tidy_tweets_SA, NUM_TOPICS_5, SA_SUFFIX)

# Global
plot_overall_topics(tidy_tweets_global, NUM_TOPICS_5, GLOBAL_SUFFIX)
```

## Overall Topics for Specific Month

```{r plot topics per month}
# past_month being the full date of the first day of that month - eg "2021-06-01" for June
plot_topics_of_month <- function(tidy_tweets_in, past_month, nr_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }

  # get the topics for the month to plot
  orgs_topics <- monthly_topics(tidy_tweets_in, past_month, nr_topics)

  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS)

  # plot top terms per month (sorted by beta relevance)
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    ggplot(aes(beta, term, fill = factor(topic))) +
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics in", format(as.Date(past_month), "%B")),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # combine plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```

```{r plot topics per month using selected number of topics (lda tuning)}
# South Africa
plot_topics_of_month(tidy_tweets_SA, "2021-01-01", NUM_TOPICS_5, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-02-01", NUM_TOPICS_4, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-03-01", NUM_TOPICS_6, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-04-01", NUM_TOPICS_5, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-05-01", NUM_TOPICS_5, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-06-01", NUM_TOPICS_6, SA_SUFFIX)

# Global
plot_topics_of_month(tidy_tweets_global, "2021-01-01", NUM_TOPICS_6, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-02-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-03-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-04-01", NUM_TOPICS_6, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-05-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-06-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
```


## Topics for Specific Media Outlet

```{r plot topics per organisation}
# org being the name of the media agency as found in the tidy_tweets dataframe
plot_topics_of_org <- function(tidy_tweets_in, org, nr_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # get the topics for the org to plot
  orgs_topics <- org_topics(tidy_tweets_in, org, nr_topics)

  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS)

  # plot top terms per media outlet (sorted by beta relevance)
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    ggplot(aes(beta, term, fill = factor(topic))) +
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics by", org),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # combine plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```

```{r plot topics per media outlet using selected number of topics (lda tuning)}
# South Africa
plot_topics_of_org(tidy_tweets_SA, SABC_NAME, NUM_TOPICS_3, SA_SUFFIX)
plot_topics_of_org(tidy_tweets_SA, DM_NAME, NUM_TOPICS_3, SA_SUFFIX)
plot_topics_of_org(tidy_tweets_SA, EWN_NAME, NUM_TOPICS_3, SA_SUFFIX)
plot_topics_of_org(tidy_tweets_SA, N24_NAME, NUM_TOPICS_5, SA_SUFFIX)

# Global
plot_topics_of_org(tidy_tweets_global, NYT_NAME, NUM_TOPICS_5, GLOBAL_SUFFIX)
plot_topics_of_org(tidy_tweets_global, ECON_NAME, NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_org(tidy_tweets_global, BBC_NAME, NUM_TOPICS_5, GLOBAL_SUFFIX)
plot_topics_of_org(tidy_tweets_global, AP_NAME, NUM_TOPICS_4, GLOBAL_SUFFIX)
```

## Topics for specific Media Outlet for specific Month

```{r plot topics per organisation per month}
# org being the name of the media agency as found in the tidy_tweets dataframe
# past_month being the full date of the first day of that month - eg "2021-06-01" for June
plot_topics_of_org_for_month <- function(tidy_tweets_in, org, past_month, nr_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # topics for org for month
  orgs_topics <- monthly_org_topics(tidy_tweets_in, org, past_month, nr_topics)
  if (is.null(orgs_topics)) {
    return()
  }
  
  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS) # force number of terms to be TOP_WORDS per topic

  # plot top terms per media outlet per month (sorted by beta relevance)
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>%
    ggplot(aes(beta, term, fill = factor(topic))) +
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics by", org, "in", format(as.Date(past_month), "%B")),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # combine plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```

```{r plot topics per outlet per month using selected number of topics (lda tuning)}
plot_process_org_per_month <- function(date, suffix) {
  # set plotting variables according to whether it is for SA or Global
  if (suffix == SA_SUFFIX) {
    outlet_list <- SA_media_outlets
    tidy_df <- tidy_tweets_SA
  } else {
    outlet_list <- global_media_outlets
    tidy_df <- tidy_tweets_global
  }
  # plot
  for (outlet in outlet_list$name) {
    print(plot_topics_of_org_for_month(tidy_df, outlet, date, NUM_TOPICS_4, suffix))
  }
}

# plot outlet topic per month
for (month in MONTH_LIST) {
  plot_process_org_per_month(month, SA_SUFFIX)
  plot_process_org_per_month(month, GLOBAL_SUFFIX)
}
```



<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Wrangling: Sentiment per topic -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

# Data Wrangling: Sentiment per topic

## Get topics into usable format
```{r split topics into list of dataframes}
# split overall topics into a list with each dataframe only containing one topic and the term associated with it
term_per_topic_list <- top_words_per_topic(overall_topics(tidy_tweets_SA, NUM_TOPICS_5), SENTIMENT_WORDS_PER_TOPIC) %>% 
  select(topic, term) %>% 
  group_split(topic)
head(term_per_topic_list)

global_term_per_topic_list <- top_words_per_topic(overall_topics(tidy_tweets_global, NUM_TOPICS_5), SENTIMENT_WORDS_PER_TOPIC) %>% 
  select(topic, term) %>% 
  group_split(topic)
head(global_term_per_topic_list)
```

```{r Create list with topics per month}
# South Africa
topics_per_month_list <- list(
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-01-01", NUM_TOPICS_5), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-02-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-03-01", NUM_TOPICS_6), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-04-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-05-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-06-01", NUM_TOPICS_6), SENTIMENT_WORDS_PER_TOPIC)
)
head(topics_per_month_list)

# Global
global_topics_per_month_list <- list(
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-01-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-02-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-03-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-04-01", NUM_TOPICS_6), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-05-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-06-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC)
)
head(global_topics_per_month_list)
```

```{r add months as an additional column to topic per months lists}
add_month <- function(df_list) {
  i <- 1
  for (month in MONTH_LIST) {
    df_list[[i]]$month <- month
    i <- i + 1
  }
  return(df_list)
}
topics_per_month_list <- add_month(topics_per_month_list)
global_topics_per_month_list <- add_month(global_topics_per_month_list)
```

```{r get topics per media outlet per month (Jan and June)}
sentiment_org_per_month <- function(date, suffix) {
  # set variables according to whether it is for SA or Global
  if (suffix == SA_SUFFIX) {
    outlet_list <- SA_media_outlets
    tidy_df <- tidy_tweets_SA
  } else {
    outlet_list <- global_media_outlets
    tidy_df <- tidy_tweets_global
  }
  
  # get topics per outlet per month
  top_words_per_topic_list <- list()
  i <- 1
  for (outlet in outlet_list$name) {
    top <- top_words_per_topic(monthly_org_topics(tidy_df, outlet, date, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC)
    top[["outlet"]] <- outlet
    top[["month"]] <- date
    top_words_per_topic_list[[i]] <- top
    i <- i + 1
  }
  return(top_words_per_topic_list)
}

# compare January and June topic sentiments per media outlet
topics_per_outlet_month_jan <- sentiment_org_per_month("2021-01-01", SA_SUFFIX)
topics_per_outlet_month_june <- sentiment_org_per_month("2021-06-01", SA_SUFFIX)
global_topics_per_outlet_month_jan <- sentiment_org_per_month("2021-01-01", GLOBAL_SUFFIX)
global_topics_per_outlet_month_june <- sentiment_org_per_month("2021-06-01", GLOBAL_SUFFIX)
```

## Get sentiment
```{r apply vader sentiment analysis to list of topics on covid tweets}
# overall topics
apply_vader <- function(df) {
  return(vader_df(df$term))
}

# topics for a specific month
apply_vader_month <- function(df) {
  vader_applied <- vader_df(df$term)
  vader_applied$topic <- df$topic
  vader_applied$month <- df$month
  vader_applied <- vader_applied %>%
    select(topic, month, compound)
  return(vader_applied)
}
  
# for individual topics
apply_vader_topic <- function(df) {
  vader_applied <- vader_df(df$term)
  vader_applied$topic <- df$topic
  vader_applied$outlet <- df$outlet
  vader_applied <- vader_applied %>%
    select(topic, outlet, compound)
  return(vader_applied)
}
```

```{r get the sentiment mean on the vader sentiment analysis}
# overall sentiment for overall topics
get_sentiment_mean <- function(df) {
  return(mean(df$compound))
}

# sentiment for topics in  a specific month
get_mean_sentiment_per_month <- function(df) {
  mean_sentiment <- df %>%
    group_by(topic, month) %>%
    summarize(compound = mean(compound))
  return(mean_sentiment)
}

# sentiment for a individual topics
get_mean_sentiment_per_topic <- function(df) {
  mean_sentiment <- df %>%
    group_by(topic, outlet) %>%
    summarize(compound = mean(compound))
  return(mean_sentiment)
}
```


```{r South Africa: get topic sentiment mean per topic}
# set names of topics
TOPIC <- list(paste0(1:length(term_per_topic_list)))

# get average sentiment
vader_analysis_per_topic <- lapply(term_per_topic_list, apply_vader)
sentiment_mean_per_topic <- unlist(lapply(vader_analysis_per_topic, get_sentiment_mean))

# get sentiment mean per topic
overall_sentiment_per_topic <- data.frame(TOPIC, sentiment_mean_per_topic)
colnames(overall_sentiment_per_topic)[1] = "topic"
head(overall_sentiment_per_topic)
```
```{r Global: get topic sentiment mean per topic}
GLOBAL_TOPIC <- list(paste0(1:length(global_term_per_topic_list)))

# get average sentiment
global_vader_analysis_per_topic <- lapply(global_term_per_topic_list, apply_vader)
global_sentiment_mean_per_topic <- unlist(lapply(global_vader_analysis_per_topic, get_sentiment_mean))

# get sentiment mean per topic
global_overall_sentiment_per_topic <- data.frame(GLOBAL_TOPIC, global_sentiment_mean_per_topic)
colnames(global_overall_sentiment_per_topic)[1] = "topic"
head(global_overall_sentiment_per_topic)
```

```{r South Africa: get topic sentiment per media outlet: Jan and June}
# get sentiment
vader_analysis_per_outlet_month_jan <- lapply(topics_per_outlet_month_jan, apply_vader_topic)
vader_analysis_per_outlet_month_june <- lapply(topics_per_outlet_month_june, apply_vader_topic)

# get sentiment mean
topic_sentiment_per_outlet_month_jan <- lapply(vader_analysis_per_outlet_month_jan, get_mean_sentiment_per_topic)
topic_sentiment_per_outlet_month_june <- lapply(vader_analysis_per_outlet_month_june, get_mean_sentiment_per_topic)
```

```{r Global: get topic sentiment per media outlet: Jan and June}
# get sentiment
global_vader_analysis_per_outlet_month_jan <- lapply(global_topics_per_outlet_month_jan, apply_vader_topic)
global_vader_analysis_per_outlet_month_june <- lapply(global_topics_per_outlet_month_june, apply_vader_topic)

# get sentiment mean
global_topic_sentiment_per_outlet_month_jan <- lapply(global_vader_analysis_per_outlet_month_jan, get_mean_sentiment_per_topic)
global_topic_sentiment_per_outlet_month_june <- lapply(global_vader_analysis_per_outlet_month_june, get_mean_sentiment_per_topic)
```

```{r South Africa: get topic sentiment per month}
# get sentiment mean
vader_analysis_per_month <- lapply(topics_per_month_list, apply_vader_month)
monthly_topic_sentiment <- lapply(vader_analysis_per_month, get_mean_sentiment_per_month)
head(monthly_topic_sentiment)
```


```{r Global: get topic sentiment per month}
# get sentiment mean
global_vader_analysis_per_month <- lapply(global_topics_per_month_list, apply_vader_month)
global_monthly_topic_sentiment <- lapply(global_vader_analysis_per_month, get_mean_sentiment_per_month)
head(global_monthly_topic_sentiment)
```


<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Visualisation: Sentiment per topic -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

# Data Visualisation: Sentiment per topic
```{r plot overall topic sentiment}
plot_sentiment_topic_bar <- function(df, plot_col, suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # plot
  ggplot(df, aes(topic, plot_col, fill=topic)) +
    geom_col() +
    labs(title = "Average sentiment per overall topic",
         subtitle = suffix,
         x="Topic",
         y="Vader Sentiment Score") +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme() + 
    theme(axis.text.x = element_text(size = 10))
}
```
```{r Plot topic sentiment per month and topic}
# South Africa
plot_sentiment_topic_bar(overall_sentiment_per_topic, sentiment_mean_per_topic, SA_SUFFIX)

# Global
plot_sentiment_topic_bar(global_overall_sentiment_per_topic, global_sentiment_mean_per_topic, GLOBAL_SUFFIX)
```


```{r plot topic sentiment per month}
plot_topic_month <- function(df, suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # plot
  ggplot() +
    geom_col(bind_rows(df, .id = "id"), mapping = aes(as.Date(month), compound, fill=factor(topic)), stat='Identity', position=position_dodge()) +
    labs(title = "Topic sentiment per month",
         subtitle = suffix,
         x=NULL,
         y="Vader Sentiment Score") +
    scale_fill_manual(values = plot_colors, name = "Topic") +
    scale_x_date(date_breaks = "1 month", date_labels = "%B") +
    custom_column_basic_theme() +
    theme(legend.position = "right", axis.text.x = element_text(size = 7.5))
}

plot_topic_month(monthly_topic_sentiment, SA_SUFFIX)
plot_topic_month(global_monthly_topic_sentiment, GLOBAL_SUFFIX)
```



```{r plot sentiment per media outlet per month: Jan and June}
plot_topic_outlet_month <- function(df, month, suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # plot
  ggplot() +
    geom_col(bind_rows(df, .id = "id"), mapping = aes(topic, compound, fill=factor(topic)), stat='Identity', position=position_dodge()) +
    labs(title = paste("Sentiment per media outlet topic in", month),
         subtitle = suffix,
         x=NULL,
         y="Vader Sentiment Score") +
    scale_fill_manual(values = plot_colors, name = "Topic") +
    facet_wrap(~ outlet, scales = "free") +
    custom_column_basic_theme()
}

plot_topic_outlet_month(topic_sentiment_per_outlet_month_jan, "January", SA_SUFFIX)
plot_topic_outlet_month(topic_sentiment_per_outlet_month_june, "June", SA_SUFFIX)
plot_topic_outlet_month(global_topic_sentiment_per_outlet_month_jan, "January", GLOBAL_SUFFIX)
plot_topic_outlet_month(global_topic_sentiment_per_outlet_month_june, "June", GLOBAL_SUFFIX)
```


