---
title: "R Notebook"
output: html_notebook
---

## adapted from TidyTextMining Ch 6

# TODO
* Try change plotting to loop maybe - changing the date?
* TODO: (nice to have) update report graphs with new x-axis title size

<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Setup -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

```{r set colours}
# Microsoft Sharepoint main colours
# https://docs.microsoft.com/en-us/sharepoint/dev/design/themes-colors
ms_red <- "#A4262C"
ms_orange <- "#CA5010"
ms_gold <- "#8F7034"
ms_green <- "#407855"
ms_teal <- "#038387"
ms_blue <- "#0078d4"
ms_darkblue <- "#40587C"
ms_indigo <- "#4052AB"
ms_plum <- "#854085"
ms_purple <- "#8764B8"
ms_coolgrey <- "#737373"
ms_warmgrey <- "#867365"

ms_colors <- c(ms_darkblue, ms_red, ms_orange, ms_teal, ms_gold, ms_coolgrey, ms_purple, ms_warmgrey, ms_green, ms_blue, ms_indigo, ms_plum)
```

```{r custom themes}
custom_column_basic_theme <- function () { 
    theme_minimal(base_size=12, base_family="Avenir") +
        theme(
            panel.background  = element_blank(),
            panel.grid.minor = element_blank(),
            legend.position="none", # no legend
            plot.title = element_text(size = 12, hjust = 0),
            plot.subtitle = element_text(size = 10, hjust = 0),
            axis.title.y = element_text(size = 10),
            axis.title.x = element_text(size = 10),
            axis.text.x = element_text(size = 6)
        )
}

custom_line_basic_theme <- function () { 
    theme_minimal(base_size=12, base_family="Avenir") +
        theme(
            panel.background  = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = 12, hjust = 0),
            plot.subtitle = element_text(size = 10, hjust = 0),
            axis.title.y = element_text(size = 10)
        )
}
```

```{r import libraries}
library(tidytext)
library(lubridate)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(topicmodels) # NOTE: (Gina) for Ubuntu need to install gsl package first with sudo apt-get install libgsl0-dev
library(ldatuning) # NOTE: (Gina) for Ubuntu needed to install (gmp package first with sudo apt-get install libgmp-dev) and (Rmpfr package with sudo apt-get install libmpfr-dev)

# Gina added
library(reshape2) # threw an error when assigning org_lda for not having this package
library(vader) # sentiment analysis
# library(data.table) # use for loops to iterate over months for plotting
```
```{r global (static) variables}
TOP_WORDS <- 10 # number of top words to show when plotting topics
TUNING <- FALSE # controls whether LDA tuning should be performed

# LDA tuning number of topic variables
NUM_TOPICS_3 <- 3
NUM_TOPICS_4 <- 4 # all outlet for month
NUM_TOPICS_5 <- 5
NUM_TOPICS_6 <- 6

# sentiment analysis variables
SENTIMENT_WORDS_PER_TOPIC <- 200
MONTHS_TO_PLOT <- seq(as.Date("2021/1/1"), by = "month", length.out = 6)

# media outlet string names
DM_NAME <- "Daily Maverick"
SABC_NAME <- "SABC News"
EWN_NAME <- "Eyewitness News"
N24_NAME <- "News24"
OUTLETS_TO_PLOT <- c(DM_NAME, SABC_NAME, EWN_NAME, N24_NAME)

NYT_NAME <- "The New York Times"
ECON_NAME <- "The Economist"
BBC_NAME <- "BBC News (World)"
AP_NAME <- "The Associated Press"
GLOBAL_OUTLETS_TO_PLOT <- c(NYT_NAME, ECON_NAME, BBC_NAME, AP_NAME)

# plotting suffixes
SA_SUFFIX <- "South Africa"
GLOBAL_SUFFIX <- "Global"
SA_GOV_SUFFIX <- "SA Government"
```



```{r extract relevant data from read in csv}
extract_relevant_data <- function(df) {
  return(df %>% select(status_id, screen_name, name, location, description, created_at, statuses_count, is_retweet, favorite_count, retweet_count, followers_count, friends_count, text))
}
```

```{r read data}
# SA media organisations
SABCNews_tweets <- extract_relevant_data(read.csv("data/sabcnews.csv"))
DailyMav_tweets <- extract_relevant_data(read.csv("data/dailymaverick.csv"))
EWN_tweets <- extract_relevant_data(read.csv("data/ewnupdates.csv"))
News24_tweets <- extract_relevant_data(read.csv("data/news24.csv"))

SA_media_outlet_list <- list(DailyMav_tweets, EWN_tweets, News24_tweets, SABCNews_tweets)

head(SA_media_outlet_list[1])

# SA government
govZA_tweets <- extract_relevant_data(read.csv("data/governmentza.csv"))

# Global
nyTimes_tweets <- extract_relevant_data(read.csv("data/nytimes.csv"))
economist_tweets <- extract_relevant_data(read.csv("data/theeconomist.csv"))
bbcworld_tweets <- extract_relevant_data(read.csv("data/bbcworld.csv"))
AP_tweets <- extract_relevant_data(read.csv("data/ap.csv"))

global_media_outlet_list <- list(nyTimes_tweets, economist_tweets, bbcworld_tweets, AP_tweets)

head(global_media_outlet_list[1])
```
<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Wrangling -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

```{r extract tweets related to covid and combine covid tweets into a list}
# extract covid tweets function
get_covid_tweets <- function(df) {
relevant_df <-  df[grepl("covid", df[["text"]]) | 
                  grepl("Covid", df[["text"]]) |
                  grepl("corona", df[["text"]]) |
                  grepl("Corona", df[["text"]]) |
                  grepl("Pandemic", df[["text"]]) |
                  grepl("pandemic", df[["text"]]), ]
relevant_df <- relevant_df %>%
  mutate(created_at = as.Date(created_at))
return(relevant_df)
}

# South African: combine relevant (Covid-related) tweets for all media outlets into list
covid_tweets_SA <- list(get_covid_tweets(DailyMav_tweets), 
                     get_covid_tweets(EWN_tweets), 
                     get_covid_tweets(News24_tweets), 
                     get_covid_tweets(SABCNews_tweets))

# SA government relevant (Covid-related) tweets
relevant_govZA <- get_covid_tweets(govZA_tweets)

# Global: combine relevant (Covid-related) tweets for all media outlets into list
covid_tweets_global <- list(get_covid_tweets(nyTimes_tweets), 
                            get_covid_tweets(AP_tweets),
                            get_covid_tweets(bbcworld_tweets), 
                            get_covid_tweets(economist_tweets))
```

```{r}
# combine all tweets into one dataframe
# Covid-related tweets
tweets_SA <- bind_rows(covid_tweets_SA)
tweets_global <- bind_rows(covid_tweets_global)
# all tweets
all_tweets_SA <- bind_rows(SA_media_outlet_list)
all_tweets_global <- bind_rows(global_media_outlet_list)
```

```{r media outlets}
SA_media_outlets <- unique(tweets_SA[c("name")])
global_media_outlets <- unique(tweets_global[c("name")])
```

```{r tidy tweets - remove stopwords and clean data}
clean_tweets <- function(tweets_in) {
  # define custom stopwords
  my_stopwords <- c("covid", "covid19", "corona", "coronavirus", "pandemic", "#sabcnews",
                    "maverick", "amp", "#covid", "#covid19", "#coronavirus",
                    "south", "africa")
  # remove stopwords
  tidy_tweets <- tweets_in %>% 
    unnest_tokens(word, text, token = "tweets") %>%
    filter(!word %in% stop_words$word, # remove normal english stopwords
           # remove normal english stopwords that would contain ' that was lost by unnesting
           # eg "ain't" is a stopwords but wouldn't be recognised as such because it was converted to "aint"
           !word %in% str_remove_all(stop_words$word, "'"), # remove remaining normal stopwords
           !word %in% my_stopwords, # remove custom stopwords
           !str_starts(word, "http"), # remove links
           !str_starts(word, "@"), # remove mentions
           str_detect(word, "^[a-z#]*$")) %>% # remove special characters
    mutate(word = gsub("\\d+", "", word)) # remove numbers

  # change "vaccines" to "vaccine" so it is treated as the same word
  tidy_tweets["word"][tidy_tweets["word"] == "vaccines"] <- "vaccine"
  return(tidy_tweets)
}

tidy_tweets_SA <- clean_tweets(tweets_SA)
tidy_tweets_global <- clean_tweets(tweets_global)
tidy_tweets_govZA <- clean_tweets(relevant_govZA)
```

<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Topic Modeling Setup -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

```{r extract top words per topic}
# topics arg: df containing word, topic and beta (probability that the word occurs in that topic) values
# words_per_topic: number of top words to get
top_words_per_topic <- function(topics, words_per_topic) {
  return(topics%>%
  group_by(topic) %>% # group words by topic
  slice_max(beta, n = words_per_topic) %>% # get most relevant words for that topic
  ungroup() %>%
  arrange(topic, -beta)) # arrange words with decreasing relevance to the topic
}
```


```{r LDA tuning}
# to find uptimal number of topics to model for given document term matrix
# taken from: https://cran.r-project.org/web/packages/ldatuning/vignettes/topics.html
lda_tuning <- function(orgs_dtm) {
  result <- FindTopicsNumber(
      orgs_dtm,
      topics = seq(from = 2, to = 8, by = 1),
      metrics = c("Griffiths2004", "CaoJuan2009", "Arun2010", "Deveaud2014"),
      method = "Gibbs",
      control = list(seed = 77),
      mc.cores = 2L,
      verbose = TRUE
    )
  # choose extremum in following plots as optimal number of topics
  FindTopicsNumber_plot(result)
}
```

# Overall topics

```{r apply topic modelling to get overall topics}
overall_topics <- function(tidy_tweets_in, nr_of_topics) {
  # word count by media organisation
  word_counts <- tidy_tweets_in %>%
    select(name, word, created_at) %>%
    count(name, word, sort = TRUE) %>% # count how often each org uses each word
    ungroup()
  
  # create document term matrix from word count
  # (treating all tweets from one organisation as one document)
  orgs_dtm <- word_counts %>%
    cast_dtm(name, word, n)
  
  # checking for ideal number of topics
  # set to TRUE to run ldatuning
  if (TUNING) {
    lda_tuning(orgs_dtm)
  }
  
  # apply topic modelling to dtm
  # orgs_lda contains all the LDA output, incl alpha, beta, gamma values, terms, documents, wordassignments etc
  orgs_lda <- LDA(orgs_dtm, k = nr_of_topics, control = list(seed = 1234))
  
  # extract relevant data from the LDA output: topic, term, beta
  orgs_topics_overall <- tidy(orgs_lda, matrix = "beta") # beta = probability that a word is in each topic
  
  return(orgs_topics_overall)
}
```

```{r plot topics}
plot_overall_topics <- function(tidy_tweets_in, nr_of_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # get the topics to plot
  orgs_topics <- overall_topics(tidy_tweets_in, nr_of_topics)
  
  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # order terms to decreasing relevance
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS) # force number of terms to be TOP_WORDS per topic
  
  # plot top terms per topic
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # reorder top terms to decreasing relevance
    ggplot(aes(beta, term, fill = factor(topic))) + # plot relevance of word for top words
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics between Jan-June 2021"),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```
 
# Overall Topics for Specific Month
 
```{r get topics per month}
monthly_topics <- function(tidy_tweets_in, past_month, nr_topics) {
  # word count by media organisation for specified month
  word_counts <- tidy_tweets_in %>%
    # create time bins (monthly)
    mutate(time_floor = floor_date(created_at, unit = "month")) %>%
    select(name, word, time_floor) %>%
    # filter to specified month
    filter(time_floor == as.Date(past_month)) %>%
    count(name, word, sort = TRUE) %>% # count how often each org uses each word
    ungroup()
    
  # create document term matrix from word count
  # (treating all tweets from one organisation as one document)
  orgs_dtm <- word_counts %>%
    cast_dtm(name, word, n)

  # checking for ideal number of topics
  # set to TRUE to run ldatuning
  if (TUNING) {
    lda_tuning(orgs_dtm)
  }
  
  # apply topic modelling to dtm
  # orgs_lda contains all the LDA output, incl alpha, beta, gamma values, terms, documents, wordassignments etc
  orgs_lda <- LDA(orgs_dtm, k = nr_topics, control = list(seed = 1234))

  # extract relevant data from the LDA output: topic, term, beta
  orgs_topics_month <- tidy(orgs_lda, matrix = "beta")
  
  return(orgs_topics_month)
}
```

```{r plot topics per month}
# past_month being the full date of the first day of that month - eg "2021-06-01" for June
plot_topics_of_month <- function(tidy_tweets_in, past_month, nr_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }

  # get the topics for the month to plot
  orgs_topics <- monthly_topics(tidy_tweets_in, past_month, nr_topics)

  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # order terms to decreasing relevance
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS) # force number of terms to be TOP_WORDS per topic

  # plot top terms per topic
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # reorder top terms to decreasing relevance
    ggplot(aes(beta, term, fill = factor(topic))) + # plot relevance of word for top words
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics in", format(as.Date(past_month), "%B")),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```

 
# Topics for Specific Media Outlet

```{r get topics per organisation}
org_topics <- function(tidy_tweets_in, org, nr_topics) {
  # word count for specified media organisation
  word_counts <- tidy_tweets_in %>%
    select(name, word) %>%
    # filter to specified org
    filter(name == org) %>%
    count(name, word, sort = TRUE) %>% # count how often the org uses each word
    ungroup()
    
  # create document term matrix
  # (treating all tweets from one organisation as one document)
  orgs_dtm <- word_counts %>%
    cast_dtm(name, word, n)
  
  # checking for ideal number of topics
  # set to TRUE to run ldatuning
  if (TUNING) {
    lda_tuning(orgs_dtm)
  }

  # apply topic modelling to dtm
  # orgs_lda contains all the LDA output, incl alpha, beta, gamma values, terms, documents, wordassignments etc
  orgs_lda <- LDA(orgs_dtm, k = nr_topics, control = list(seed = 1234))

  # extract relevant data from the LDA output: topic, term, beta
  org_topics <- tidy(orgs_lda, matrix = "beta")
  
  return(org_topics)
}
```

```{r plot topics per organisation}
# org being the name of the media agency as found in the tidy_tweets dataframe
plot_topics_of_org <- function(tidy_tweets_in, org, nr_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # get the topics for the org to plot
  orgs_topics <- org_topics(tidy_tweets_in, org, nr_topics)

  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # order terms to decreasing relevance
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS) # force number of terms to be TOP_WORDS per topic

  # plot top terms per topic
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # reorder top terms to decreasing relevance
    ggplot(aes(beta, term, fill = factor(topic))) + # plot relevance of word for top words
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics by", org),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```


# Topics for specific Media Outlet for specific Month

```{r get topics per organisation per month}
monthly_org_topics <- function(tidy_tweets_in, org, past_month, nr_topics) {
  # word count for specified media organisation
  word_counts <- tidy_tweets_in %>%
    mutate(time_floor = floor_date(created_at, unit = "month")) %>%
    select(name, word, time_floor) %>%
    # filter to specified org
    filter(name == org) %>%
    # filter to specified month
    filter(time_floor == as.Date(past_month)) %>%
    count(name, word, sort = TRUE) %>% # count how often the org uses each word
    ungroup()
  
  if (length(word_counts$name) == 0) {
    print(paste(org, "did not tweet the month starting on", past_month))
    return()
  }
    
  # create document term matrix
  # (treating all tweets from one organisation as one document)
  orgs_dtm <- word_counts %>%
    cast_dtm(name, word, n)
  
  # checking for ideal number of topics
  # set to TRUE to run ldatuning
  if (TUNING) {
    lda_tuning(orgs_dtm)
  }

  # apply topic modelling to dtm
  # orgs_lda contains all the LDA output, incl alpha, beta, gamma values, terms, documents, wordassignments etc
  orgs_lda <- LDA(orgs_dtm, k = nr_topics, control = list(seed = 1234))

  # extract relevant data from the LDA output: topic, term, beta
  topics <- tidy(orgs_lda, matrix = "beta")
  return(topics)
}
```

```{r plot topics per organisation per month}
# org being the name of the media agency as found in the tidy_tweets dataframe
# past_month being the full date of the first day of that month - eg "2021-06-01" for June
plot_topics_of_org_for_month <- function(tidy_tweets_in, org, past_month, nr_topics, title_suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (title_suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  # topics for org for month
  orgs_topics <- monthly_org_topics(tidy_tweets_in, org, past_month, nr_topics)
  if (is.null(orgs_topics)) {
    return()
  }
  
  # top words per topic
  top_terms <- top_words_per_topic(orgs_topics, TOP_WORDS) %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # order terms to decreasing relevance
    group_by(topic) %>%
    filter(row_number()<=TOP_WORDS) # force number of terms to be TOP_WORDS per topic

  # plot top terms per topic
  top_terms %>%
    mutate(term = reorder_within(term, beta, topic)) %>% # reorder top terms to decreasing relevance
    ggplot(aes(beta, term, fill = factor(topic))) + # plot relevance of word for top words
    geom_col(show.legend = FALSE) +
    labs(title = paste("Covid-related Topics by", org, "in", format(as.Date(past_month), "%B")),
         subtitle = title_suffix) +
    facet_wrap(~ topic, scales = "free") + # plot for each topic
    scale_y_reordered() +
    scale_fill_manual(values = plot_colors) +
    custom_column_basic_theme()
}
```


<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Visualisation -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

# Overall topics

```{r}
# South Africa
plot_overall_topics(tidy_tweets_SA, NUM_TOPICS_6, SA_SUFFIX)
plot_overall_topics(tidy_tweets_govZA, NUM_TOPICS_6, SA_GOV_SUFFIX)
# Global
plot_overall_topics(tidy_tweets_global, NUM_TOPICS_5, GLOBAL_SUFFIX)
```

# Overall Topics for Specific Month

```{r}
# South Africa

plot_topics_of_month(tidy_tweets_SA, "2021-01-01", NUM_TOPICS_5, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-02-01", NUM_TOPICS_4, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-03-01", NUM_TOPICS_6, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-04-01", NUM_TOPICS_5, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-05-01", NUM_TOPICS_5, SA_SUFFIX)
plot_topics_of_month(tidy_tweets_SA, "2021-06-01", NUM_TOPICS_6, SA_SUFFIX)
```


```{r}
# Global

plot_topics_of_month(tidy_tweets_global, "2021-01-01", NUM_TOPICS_6, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-02-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-03-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-04-01", NUM_TOPICS_6, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-05-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_month(tidy_tweets_global, "2021-06-01", NUM_TOPICS_4, GLOBAL_SUFFIX)
```


# Topics for Specific Media Outlet

```{r}
# South Africa

plot_topics_of_org(tidy_tweets_SA, "SABC News", NUM_TOPICS_3, SA_SUFFIX)
plot_topics_of_org(tidy_tweets_SA, "Daily Maverick", NUM_TOPICS_3, SA_SUFFIX)
plot_topics_of_org(tidy_tweets_SA, "Eyewitness News", NUM_TOPICS_3, SA_SUFFIX)
plot_topics_of_org(tidy_tweets_SA, "News24", NUM_TOPICS_5, SA_SUFFIX)
```

```{r}
# Global

plot_topics_of_org(tidy_tweets_global, "The New York Times", NUM_TOPICS_5, GLOBAL_SUFFIX)
plot_topics_of_org(tidy_tweets_global, "The Economist", NUM_TOPICS_4, GLOBAL_SUFFIX)
plot_topics_of_org(tidy_tweets_global, "BBC News (World)", NUM_TOPICS_5, GLOBAL_SUFFIX)
plot_topics_of_org(tidy_tweets_global, "The Associated Press", NUM_TOPICS_4, GLOBAL_SUFFIX)
```

# Topics for specific Media Outlet for specific Month

```{r plot topics per outlet per month}
plot_process_org_per_month <- function(date, suffix) {
  # set plotting variables according to whether it is for SA or Global
  if (suffix == SA_SUFFIX) {
    outlet_list <- SA_media_outlets
    tidy_df <- tidy_tweets_SA
  } else {
    outlet_list <- global_media_outlets
    tidy_df <- tidy_tweets_global
  }
  # plot
  for (outlet in outlet_list$name) {
    print(plot_topics_of_org_for_month(tidy_df, outlet, date, NUM_TOPICS_4, suffix))
  }
}

# plot outlet topic per month
month_list <- c("2021-01-01", "2021-02-01", "2021-03-01", "2021-04-01", "2021-05-01", "2021-06-01")
for (month in month_list) {
  plot_process_org_per_month(month, SA_SUFFIX)
  plot_process_org_per_month(month, GLOBAL_SUFFIX)
}
```



<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Wrangling: Sentiment per topic -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->

# Sentiment per topic

```{r split topics into list of dataframes}
# split overall topics into a list with each dataframe only containing one topic and the term associated with it
term_per_topic_list <- top_words_per_topic(overall_topics(tidy_tweets_SA, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC) %>% 
  select(topic, term) %>% 
  group_split(topic)
head(term_per_topic_list)

global_term_per_topic_list <- top_words_per_topic(overall_topics(tidy_tweets_global, NUM_TOPICS_5), SENTIMENT_WORDS_PER_TOPIC) %>% 
  select(topic, term) %>% 
  group_split(topic)
head(global_term_per_topic_list)
```

```{r Create list with topics per month}
# TODO: optimise this somehow by passing tidy tweets, num_topics and date
# TODO: Should indicate date on dataframe such that we know which month is associated with which dataframe in topics_per_month
# TODO: optimise this in some way with a loop?

# South Africa: topics per month
topics_per_month_list <- list(
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-01-01", NUM_TOPICS_5), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-02-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-03-01", NUM_TOPICS_6), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-04-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-05-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_SA, "2021-06-01", NUM_TOPICS_6), SENTIMENT_WORDS_PER_TOPIC)
)
head(topics_per_month_list)

# Global: topics per month
global_topics_per_month_list <- list(
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-01-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-02-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-03-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-04-01", NUM_TOPICS_6), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-05-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_topics(tidy_tweets_global, "2021-06-01", NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC)
)
head(global_topics_per_month_list)
```



```{r Create list with topics per media outlet}
# TODO: optimise this somehow by passing tidy tweet, num_topics and media outlet name
# TODO: Should indicate date on dataframe such that we know which month is associated with which dataframe in topics_per_outlet
# TODO: optimise this in some way with a loop?

# South Africa: topics per outlet
topics_per_outlet_list <- list(
  top_words_per_topic(org_topics(tidy_tweets_SA, DM_NAME, NUM_TOPICS_3), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(org_topics(tidy_tweets_SA, SABC_NAME, NUM_TOPICS_3), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(org_topics(tidy_tweets_SA, EWN_NAME, NUM_TOPICS_3), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(org_topics(tidy_tweets_SA, N24_NAME, NUM_TOPICS_5), SENTIMENT_WORDS_PER_TOPIC)
)
head(topics_per_outlet_list)

# Global: topics per outlet
global_topics_per_outlet_list <- list(
  top_words_per_topic(org_topics(tidy_tweets_global, NYT_NAME, NUM_TOPICS_3), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(org_topics(tidy_tweets_global, ECON_NAME, NUM_TOPICS_3), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(org_topics(tidy_tweets_global, BBC_NAME, NUM_TOPICS_5), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(org_topics(tidy_tweets_global, AP_NAME, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC)
)
head(global_topics_per_outlet_list)
```

```{r Create list with topics per media outlet per month}
# TODO: optimise this somehow by passing tidy tweet, date, and media outlet name
# TODO: add Jan and figure out how to group bar plot
JUNE1 <- "2021-06-01"

# TODO: Should indicate date on dataframe such that we know which month is associated with which dataframe in topics_per_outlet_month_list
# TODO: optimise this in some way with a loop?

# South Africa: topics per outlet per month
topics_per_outlet_month_list <- list(
  top_words_per_topic(monthly_org_topics(tidy_tweets_SA, DM_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_org_topics(tidy_tweets_SA, SABC_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_org_topics(tidy_tweets_SA, EWN_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_org_topics(tidy_tweets_SA, N24_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC)
)
head(topics_per_outlet_month_list)

# Global: topics per outlet per month
global_topics_per_outlet_month_list <- list(
  top_words_per_topic(monthly_org_topics(tidy_tweets_global, NYT_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_org_topics(tidy_tweets_global, ECON_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_org_topics(tidy_tweets_global, BBC_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC),
  top_words_per_topic(monthly_org_topics(tidy_tweets_global, AP_NAME, JUNE1, NUM_TOPICS_4), SENTIMENT_WORDS_PER_TOPIC)
)
head(global_topics_per_outlet_month_list)
```


```{r apply vader sentiment analysis to list of topics on covid tweets}
# https://towardsdatascience.com/sentimental-analysis-using-vader-a3415fef7664
# all values greater than zeroes will be considered a positive review and all values less than zero would be considered as a negative review.
apply_vader <- function(df) {
  return(vader_df(df$term))
}

get_sentiment_mean <- function(df) {
  return(mean(df$compound))
}
```

```{r South Africa: get topic sentiment mean per topic}
# TODO need to plot nicely (as table?)

# set names of topics
TOPIC <- list(paste0(1:length(term_per_topic_list)))

# get average sentiment
vader_analysis_per_topic <- lapply(term_per_topic_list, apply_vader)
sentiment_mean_per_topic <- unlist(lapply(vader_analysis_per_topic, get_sentiment_mean))

# get sentiment mean per topic
overall_sentiment_per_topic <- data.frame(TOPIC, sentiment_mean_per_topic)
colnames(overall_sentiment_per_topic)[1] = "topic"
head(overall_sentiment_per_topic)
```
```{r Global: get topic sentiment mean per topic}
# TODO need to plot nicely (as table?)

# set names of topics
GLOBAL_TOPIC <- list(paste0(1:length(global_term_per_topic_list)))

# get average sentiment
global_vader_analysis_per_topic <- lapply(global_term_per_topic_list, apply_vader)
global_sentiment_mean_per_topic <- unlist(lapply(global_vader_analysis_per_topic, get_sentiment_mean))

# get sentiment mean per topic
global_overall_sentiment_per_topic <- data.frame(GLOBAL_TOPIC, global_sentiment_mean_per_topic)
colnames(global_overall_sentiment_per_topic)[1] = "topic"
head(global_overall_sentiment_per_topic)
```

```{r South Africa: get topic sentiment mean per month, outlet, and month per outlet}
# TODO need to plot nicely (as table?)

# get average sentiment
vader_analysis_per_month <- lapply(topics_per_month_list, apply_vader)
sentiment_mean_per_month <- unlist(lapply(vader_analysis_per_month, get_sentiment_mean))

vader_analysis_per_outlet <- lapply(topics_per_outlet_list, apply_vader)
sentiment_mean_per_outlet <- unlist(lapply(vader_analysis_per_outlet, get_sentiment_mean))

vader_analysis_per_outlet_month <- lapply(topics_per_outlet_month_list, apply_vader)
sentiment_mean_per_outlet_month <- unlist(lapply(vader_analysis_per_outlet_month, get_sentiment_mean))

# get sentiment mean per  month
overall_sentiment_per_month <- data.frame(month=MONTHS_TO_PLOT, sentiment_mean_per_month)
head(overall_sentiment_per_month)

overall_sentiment_per_outlet <- data.frame(media_outlet=OUTLETS_TO_PLOT, sentiment_mean_per_outlet)
head(overall_sentiment_per_outlet)

overall_sentiment_per_outlet_month <- data.frame(media_outlet=OUTLETS_TO_PLOT, sentiment_mean_per_outlet_month)
head(overall_sentiment_per_outlet_month)
```
```{r Global: get topic sentiment mean per month, outlet, and month per outlet}
# TODO need to plot nicely (as table?)

# get average sentiment
global_vader_analysis_per_month <- lapply(global_topics_per_month_list, apply_vader)
global_sentiment_mean_per_month <- unlist(lapply(global_vader_analysis_per_month, get_sentiment_mean))

global_vader_analysis_per_outlet <- lapply(global_topics_per_outlet_list, apply_vader)
global_sentiment_mean_per_outlet <- unlist(lapply(global_vader_analysis_per_outlet, get_sentiment_mean))

global_vader_analysis_per_outlet_month <- lapply(global_topics_per_outlet_month_list, apply_vader)
global_sentiment_mean_per_outlet_month <- unlist(lapply(global_vader_analysis_per_outlet_month, get_sentiment_mean))

# get sentiment mean per  month
global_overall_sentiment_per_month <- data.frame(month=MONTHS_TO_PLOT, global_sentiment_mean_per_month)
head(global_overall_sentiment_per_month)

global_overall_sentiment_per_outlet <- data.frame(media_outlet=OUTLETS_TO_PLOT, global_sentiment_mean_per_outlet)
head(global_overall_sentiment_per_outlet)

global_overall_sentiment_per_outlet_month <- data.frame(media_outlet=OUTLETS_TO_PLOT, global_sentiment_mean_per_outlet_month)
head(global_overall_sentiment_per_outlet_month)
```


<!----------------------------------------------------------------------------------------------------------------------------------------------------->
<!-- Data Visualisation: Sentiment per topic -->
<!----------------------------------------------------------------------------------------------------------------------------------------------------->
# TODO: 
* plot global stuff too
* plot with new data stuff

```{r bar plot function template}
plot_sentiment_topic_bar <- function(df, plot_row, plot_col, title, suffix) {
  # set plot colors
  plot_colors <- ms_colors
  if (suffix == GLOBAL_SUFFIX) {
    plot_colors <- rev(ms_colors)
  }
  
  fill = plot_row
  # set fill to month if plotting by month
  if (title == "month") {
    fill = format(plot_row, "%B")
  }
  
  x_axis_title <- NULL
  # set x-axis title to Topic if plotting by topic
  if (title == "topic") {
    x_axis_title <- "Topic"
  }
  
  # plot
ggplot(df, aes(plot_row, plot_col, fill=fill)) +
  geom_col() +
  labs(title = paste("Average sentiment per", title),
       subtitle = suffix,
       x=x_axis_title,
       y="Vader Sentiment Score") +
  scale_fill_manual(values = plot_colors) +
  custom_column_basic_theme() + 
  theme(axis.text.x = element_text(size = 10))
}
```

```{r South Africa: plot topic sentiment}
plot_sentiment_topic_bar(overall_sentiment_per_topic, overall_sentiment_per_topic$topic, sentiment_mean_per_topic, "topic", SA_SUFFIX)
plot_sentiment_topic_bar(overall_sentiment_per_month, overall_sentiment_per_month$month, sentiment_mean_per_month, "month", SA_SUFFIX)
plot_sentiment_topic_bar(overall_sentiment_per_outlet, overall_sentiment_per_outlet$media_outlet, sentiment_mean_per_outlet, "outlet", SA_SUFFIX)
plot_sentiment_topic_bar(overall_sentiment_per_outlet_month, overall_sentiment_per_outlet_month$media_outlet, sentiment_mean_per_outlet_month, "outlet (June)", SA_SUFFIX)
```
```{r Global: plot topic sentiment}
plot_sentiment_topic_bar(global_overall_sentiment_per_topic, global_overall_sentiment_per_topic$topic, global_sentiment_mean_per_topic, "topic", GLOBAL_SUFFIX)
plot_sentiment_topic_bar(global_overall_sentiment_per_month, global_overall_sentiment_per_month$month, global_sentiment_mean_per_month, "month", GLOBAL_SUFFIX)
plot_sentiment_topic_bar(global_overall_sentiment_per_outlet, global_overall_sentiment_per_outlet$media_outlet, global_sentiment_mean_per_outlet, "outlet", GLOBAL_SUFFIX)
plot_sentiment_topic_bar(global_overall_sentiment_per_outlet_month, global_overall_sentiment_per_outlet_month$media_outlet, global_sentiment_mean_per_outlet_month, "outlet (June)", GLOBAL_SUFFIX)
```
