---
title: "Assignment 1: Gina Graphs"
output: html_notebook
---

# TODO: functions to create
- max in a column
- last element in column minus second last element in column

# Import libraries
```{r}
library(tidyverse)
library(lubridate) # dates
library(formattable) # round numbers
library(ggplot2)
library(scales)
```

# Read csv
```{r}
# general (cumulative) information regarding tests, deaths, and recoveries. Complete timeline.
tests_deaths_recoveries_data <- read.csv("data/covid19za_timeline_testing.csv")
head(tests_deaths_recoveries_data)

# total (cumulative) number of confirmed cases by province
provincial_total_confirmed <- read.csv("data/covid19za_provincial_cumulative_timeline_confirmed.csv")
head(provincial_total_confirmed)

# big one
# big_data <- read.csv("data/owid-covid-data_SouthAfricaExtracted_big.csv")
# head(big_data)
```


# Deaths vs Recovered
```{r}
# total deaths
total_deaths <- max(tests_deaths_recoveries_data$deaths, na.rm = TRUE)
head(total_deaths)

# get last day's number of deaths increase
last_death_cols <- tail(tests_deaths_recoveries_data$deaths, n=2)
death_day_update <- last_death_cols[2] - last_death_cols[1]
head(death_day_update)

# total recovered
total_recovered <- max(tests_deaths_recoveries_data$recovered, na.rm = TRUE)
head(total_recovered)

# percentage recovered
percentage_recovered <- total_recovered / (total_recovered + total_deaths) * 100
head(formattable(percentage_recovered, digits = 1, format = "f")) # round to 1 digit
```
# Confirmed vs Active Total
```{r}
# total confirmed cases
total_confirmed <- max(provincial_total_confirmed$total, na.rm = TRUE)
head(total_confirmed)

# get last day's number of confirmed increase
last_confirmed_cols <- tail(provincial_total_confirmed$total, n=2)
confirmed_day_update <- last_confirmed_cols[2] - last_confirmed_cols[1]
head(confirmed_day_update)

# active cases = confirmed cases  - recoveries - deaths
total_active <- total_confirmed - total_recovered - total_deaths
head(total_active)
```

# Total number of tests and test day increase
```{r}
# get number of tests
total_tests <- max(tests_deaths_recoveries_data$cumulative_tests, na.rm = TRUE)
head(total_tests)
# get last day's number of tests increase
last_tests_cols <- tail(tests_deaths_recoveries_data$cumulative_tests, n=2)
test_day_update <- last_tests_cols[2] - last_tests_cols[1]
test_day_update
```

# Daily confirmed cases
final TODO: 
* Add lockdown levels
* line plot of 7-day rolling average (and add title)
* Format dates and values
* format y-axis integer representation

<!-- # add rolling average to the data -->
<!-- testing <- testing %>% -->
<!--   mutate(rolling_avg = roll_mean(daily_tests, n=7, fill = NA, align = "right")) -->
```{r}
confirmed <- provincial_total_confirmed %>% select(date, total) %>% mutate(date = dmy(date))

confirmed_summary <- provincial_total_confirmed %>%
  select(date, total) %>%
  mutate(date = dmy(date)) %>%
  mutate(daily_confirmed=total-lag(total,default=0))
names(confirmed_summary)[names(confirmed_summary) == 'total'] <- 'total_confirmed'

ggplot(confirmed_summary, aes(date, daily_confirmed)) + 
  geom_col() +
  labs(title= "Daily confirmed cases",
       caption = "Updated: June 2021",
       x = NULL,
       y = NULL) +
  theme_minimal()
```


# Cases vs Active cases
`active = total_confirmed - total_recovered - total_deaths`

## Get death cases
From Johanna's code (TODO remove this duplicate when merging code)
```{r Data: deaths}
death_data <- read.csv("data/covid19za_provincial_cumulative_timeline_deaths.csv")
head(death_data)
```
```{r tidying}
deaths <- death_data %>% pivot_longer(c('EC', 'FS', 'GP', 'KZN', 'LP', 'MP', 'NC', 'NW', 'WC', 'UNKNOWN'), names_to = "province", values_to = "deaths")
deaths <- deaths %>%
  select(date, province, deaths, total)  %>%
  filter(province != 'UNKNOWN') %>%
  mutate(date = dmy(date))

head(deaths)
```
### Create deaths summary
```{r total and daily deaths}
death_summary <- death_data %>%
  select(date, total) %>%
  mutate(date = dmy(date)) %>%
  mutate(daily_death=total-lag(total,default=0))
names(death_summary)[names(death_summary) == 'total'] <- 'total_deaths'
head(death_summary)
```
## Get recovery data
```{r}
recovery_summary <- tests_deaths_recoveries_data %>%
  select(date, recovered) %>%
  mutate(date = dmy(date)) %>%
  mutate(daily_recovered=recovered-lag(recovered,default=0))
names(recovery_summary)[names(recovery_summary) == 'recovered'] <- 'total_recovered'

head(recovery_summary)
```


## Join recoveries, deaths and confirmed
```{r}
# inner_join() joins based on matching keys (we only want overlapping dates)
active_df <- recovery_summary %>% 
  inner_join(death_summary, by = "date") %>% 
  inner_join(confirmed_summary, by = "date")
head(active_cases_df)
```



## Get active cases
```{r}
active_df$total_active <- with(active_df, total_confirmed - total_deaths - total_recovered)
active_df
```

### Get active cases summary
```{r}
active_confirmed_summary <- active_df %>%
  select(date, total_active, total_confirmed, daily_confirmed) %>%
  mutate(daily_active=total_active-lag(total_active,default=0))
head(active_confirmed_summary)
```


## Plot Active vs confirmed
final TODO: 
* add total active and total confirmed label
* Remove legend title
* Adjust dates shown (values and format)
* Adjust y-axis number formatting
* Gap in active cases
```{r}

ggplot(active_confirmed_summary, aes(date)) + 
  geom_line(aes(y = total_active, colour = "Active Cases")) + 
  geom_line(aes(y = total_confirmed, colour = "Confirmed Cases")) +
  labs(title= "Cases vs Active cases",
     caption = "Updated: June 2021",
     x = NULL,
     y = NULL) +
  theme_minimal()
```


